---
title: 'KV Analysis for Performing-Dynamics performances'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

***KV values from this analysis were used.***

Description: This is a rough overview of performing performance (KV) from [music-teaching (experiment2)](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-2).

Performing-Dynamics performances: data from the performing condition (dynamics)

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r extract, include = FALSE}
# onset
dt_kv <- fread("../preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv

# SubNr as a factor
dt_kv$SubNr <- as.factor(dt_kv$SubNr)

# assign RowNr
dt_kv$RowNr <- rep(1:72, nrow(dt_kv)/72)

# include only dynamics
dt_kv <- dt_kv[Condition == "performing" & Skill == "dynamics"]
dt_kv <- dt_kv[Error == 0]
```

```{r Velocitys, include = FALSE}
# define subcomponents
# for intervals
ls_legato <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_staccato <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))
ls_forte <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_piano <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))

# for each note (duration/velocity)
ls_legato2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_staccato2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))
ls_forte2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_piano2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))

# define Component Change (LtoS, FtoP)
change_1 <- c(5, 17, 47)
# define Component Change (StoL, PtoF)
change_2 <- c(8, 20, 39)

# assign subcomponents
dt_kv$Subcomponent <- "NA"

# for forte
for (phrase in 1:length(ls_forte2)){
  for (note in 1:length(ls_forte2[[phrase]])){
    dt_kv[RowNr == ls_forte2[[phrase]][note]]$Subcomponent <- "Forte"
  }
}

# for piano
for (phrase in 1:length(ls_piano2)){
  for (note in 1:length(ls_piano2[[phrase]])){
    dt_kv[RowNr == ls_piano2[[phrase]][note]]$Subcomponent <- "Piano"
  }
}
```

# KV
## 1. All - individual
- Average KV for each participant.

```{r kv, echo = FALSE}
# For each individual
kv <- dt_kv[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = length(Velocity), Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Subcomponent)]
print(kv)
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kv_bar,  echo = FALSE}
p_kv <- ggplot(data = kv, aes(x = SubNr, y = Mean)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "Key Velocity", subtitle = "KV: Performing-Dynamics")
p_kv
```

#### Sequence plot
- Right wrap: Subject Number

#### All
```{r kv_seq, fig.width = 14, fig.height = 12, echo = FALSE}
# for each individual
dt_kv$TrialNr <- as.factor(dt_kv$TrialNr)
kv_seq <- dt_kv[Subcomponent != "NA", .(N = length(Velocity), Mean = mean(Velocity)), .(SubNr, TrialNr, RowNr)]

p_kv_seq <- ggplot(data = kv_seq, aes(x = RowNr, y = Mean, group = TrialNr, colour = TrialNr)) +
  geom_line() +
  geom_point() +
  facet_grid(SubNr ~ .) +
  labs(x = "RowNr", y = "Velocity", subtitle = "KV: Performing-Dynamics(All)") +
  scale_x_continuous(breaks=seq(1,72,1))
p_kv_seq
```

#### Forte
```{r kv_seq_for, fig.width = 14, fig.height = 12, echo = FALSE}
# For each individual
dt_kv$TrialNr <- as.factor(dt_kv$TrialNr)
kv_seq_for <- dt_kv[Subcomponent == "Forte", .(N = length(Velocity), Mean = mean(Velocity)), by = .(SubNr, TrialNr, RowNr)]

p_kv_seq_for <- ggplot(data = kv_seq_for, aes(x = RowNr, y = Mean, group = TrialNr, colour = TrialNr)) +
  geom_line() +
  geom_point() +
  facet_grid(SubNr ~ .) +
  labs(x = "RowNr", y = "Key Velocity") + scale_x_continuous(breaks=seq(1,72,1))
p_kv_seq_for
```

#### Piano
```{r kv_seq_pia, fig.width = 14, fig.height = 12, echo = FALSE}
# For each individual
dt_kv$TrialNr <- as.factor(dt_kv$TrialNr)
kv_seq_pia <- dt_kv[Subcomponent == "Piano", .(N = length(Velocity), Mean = mean(Velocity)), by = .(SubNr, TrialNr, RowNr)]

p_kv_seq_pia <- ggplot(data = kv_seq_pia, aes(x = RowNr, y = Mean, group = TrialNr, colour = TrialNr)) +
  geom_line() +
  geom_point() +
  facet_grid(SubNr ~ .) +
  labs(x = "RowNr", y = "Key Velocity") + scale_x_continuous(breaks=seq(1,72,1))
p_kv_seq_pia
```

#### Distribution plot
- Right wrap: Trial Number
- Upper wrap: Subject Number

#### Forte
```{r kv_dis_for, fig.width = 14, fig.height = 8, echo = FALSE}
p_kv_dis_for <- ggplot(dt_kv[Subcomponent == "Forte"], aes(x = Velocity)) +
  geom_histogram(binwidth = 2, alpha = .5, position = "identity") +
  facet_grid(TrialNr ~ SubNr) +
  labs(x = "Key Velocity", y = "Count", subtitle = "KV: Performing-Dynamics(Forte)")
p_kv_dis_for
```

#### Piano
```{r kv_dis_pia, fig.width = 14, fig.height = 8, echo = FALSE}
p_kv_dis_pia <- ggplot(dt_kv[Subcomponent == "Piano"], aes(x = Velocity)) +
  geom_histogram(binwidth = 2, alpha = .5, position = "identity") +
  facet_grid(TrialNr ~ SubNr) +
  labs(x = "Key Velocity", y = "Count", subtitle = "KV: Performing-Dynamics(Piano)")
p_kv_dis_pia
```

## 2. All - group
- Average Velocity

```{r kv_all, echo = FALSE}
# Group mean
kv_all <- kv[, .(N = length(Mean), Mean = mean(Mean), SD = sd(Mean)), by = .(Subcomponent)]
print(kv_all)
```

# Included performances for averaging
## Forte
```{r outlier_for, echo = FALSE}
# for forte
kv_trial_for <- dt_kv[Subcomponent == "Forte", .(N = length(Velocity), Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr)]
kv_trial_for
```

## Piano
```{r outlier_pia, echo = FALSE}
# for piano
kv_trial_pia <- dt_kv[Subcomponent == "Piano", .(N = length(Velocity), Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr)]
kv_trial_pia
```

```{r export, include = FALSE}
kv_valid <- dt_kv[, .(N = length(Velocity), Mean = mean(Velocity)), .(SubNr, TrialNr)]
fwrite(kv_valid, "./kv_valid.csv")
```
