---
title: 'KV Analysis for Non-Expressive performances'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a rough overview of non-expressive performance (KV) from teaching-v2.0.

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r extract, include = FALSE}
dt_kv <- fread("./preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv

# SubNr as a factor
dt_kv$SubNr <- as.factor(dt_kv$SubNr)

# assign RowNr
dt_kv$RowNr <- rep(1:72, nrow(dt_kv)/72)
```

```{r Velocitys, include = FALSE}
# # convert bpm to ms
# dt_kv[Tempo == 120]$Tempo <- 250
# dt_kv[Tempo == 110]$Tempo <- 273
# dt_kv[Tempo == 100]$Tempo <- 300

# remove irrelevant intervals
ls_irrelevant <- list(c(27:28), c(33:35), 52, 57, c(66:72))
dt_kv$EighthNote <- "Yes"

for (phrase in 1:length(ls_irrelevant)){
  for (note in 1:length(ls_irrelevant[[phrase]])){
    dt_kv[RowNr == ls_irrelevant[[phrase]][note]]$EighthNote <- "No"
  }
}
```

# KV
## 1. All - individual
- Average KV for each participant.

```{r kv, echo = FALSE}
# For each individual
kv <- dt_kv[EighthNote == "Yes", .(N = length(Velocity), Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr)]
print(kv)
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kv_bar,  echo = FALSE}
p_kv <- ggplot(data = kv, aes(x = SubNr, y = Mean)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  labs(y = "Key Velocity", subtitle = "KV: Non-Expression")
p_kv
```

#### Sequence plot
- Right wrap: Subject Number

```{r kv_seq, fig.width = 14, fig.height = 12, echo = FALSE}
# For each individual
kv_seq <- dt_kv[EighthNote == "Yes", .(N = length(Velocity), Mean = mean(Velocity)), by = .(SubNr, TrialNr, RowNr)]

p_kv_seq <- ggplot(data = kv_seq, aes(x = RowNr, y = Mean, group = TrialNr)) +
  geom_line() +
  geom_point() +
  facet_grid(SubNr ~ .) +
  labs(x = "RowNr", y = "Key Velocity") + scale_x_continuous(breaks=seq(1,72,1))
p_kv_seq
```

#### Distribution plot
- Right wrap: Trial Number
- Upper wrap: Subject Number

```{r kv_dis, fig.width = 14, fig.height = 8, echo = FALSE}
p_kv_dis <- ggplot(dt_kv[EighthNote == "Yes"], aes(x = Velocity)) +
  geom_histogram(binwidth = 2, alpha = .5, position = "identity") +
  facet_grid(TrialNr ~ SubNr) +
  labs(x = "Key Velocity", y = "Count", subtitle = "KV: Non-Expression")
p_kv_dis
```

## 2. All - group
- Average Velocity

```{r kv_all, echo = FALSE}
# Group mean
kv_all <- kv[, .(N = length(Mean), Mean = mean(Mean), SD = sd(Mean))]
print(kv_all)
```

### Valid performances (exclude outliers)
```{r outlier, echo = FALSE}
excluded_kv <- kv[Mean < kv_all$Mean+3*kv_all$SD & Mean > kv_all$Mean-3*kv_all$SD]
print(excluded_kv)
```
- The number of excluded performance: `r kv_all$N-nrow(excluded_kv)`

# KV Variability
## 1. All - individual
- Average CV for each participant

```{r cv, echo = FALSE}
# for each participant
cv <- dt_kv[EighthNote == "Yes", .(N = length(Velocity), CV = sd(Velocity)/mean(Velocity)) , by = .(SubNr, TrialNr)]
cv
```

```{r cv_box, echo = FALSE}
p_cv_box <- ggplot(cv, aes(x = SubNr, y = CV)) +
  geom_boxplot() +
  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5)
p_cv_box
```

## 2. All - group
- Average CV

```{r cv_all, echo = FALSE}
# for all
cv_all <- cv[, .(N = length(CV), Mean = mean(CV), SD = sd(CV))]
print(cv_all)
```

### Valid performances (exclude outliers)
```{r cv_outlier, echo = FALSE}
excluded_cv <- cv[CV < cv_all$Mean+3*(cv_all$SD) & CV > cv_all$Mean-3*cv_all$SD]
print(excluded_cv)
```
- The number of excluded performance: `r cv_all$N-nrow(excluded_cv)`

```{r export, include = FALSE}
fwrite(excluded_cv, "./kv_valid.csv")
```
