---
title: 'Demographics'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of demographic information

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
```

# Questionnaire

```{r file, include = FALSE}
filename_q = "./questionnaire.csv"
```

```{r extract, echo = FALSE}
# read csv files
dt <- fread(filename_q, header = T, sep = ",", dec = ".", na.string = "NA")
dt <- dt[-c(1,2),] # exclude test/pilot data

# change some colnames
colnames(dt)[2] <- "SubNr"
colnames(dt)[3] <- "Age"
colnames(dt)[4] <- "Gender"
colnames(dt)[7] <- "Handedness"
colnames(dt)[15] <- "PianoTotalPractice"
colnames(dt)[25] <- "TeachingPiano" # there are only people who have taught the piano
colnames(dt)[27] <- "TeachingPianoYears"

dt$TeachingPianoYears <- c(7, 1, 12, 5, 30, NA, 11, 2, 5, 3, 6, NA, 3, 6, 8, 9, 3, NA, 5, 3)

# change some characteristics
dt$Age <- as.numeric(dt$Age)
dt$PianoTotalPractice <- as.numeric(dt$PianoTotalPractice)

# exclude participants
dt_included <- dt

print(dt_included)
```

## Descriptive stats for included participants
Note: A different questionnaire was used for only SubNr 21 and adjusted for necessary columns (but not others; see original csv for the precise information)

- Number of included participants: `r nrow(dt_included)` (mean age: `r round(mean(dt_included$Age), 2)`, sd: `r round(sd(dt_included$Age), 2)`)
- Male: `r nrow(dt_included[Gender == "Male"])` (mean age: `r round(mean(dt_included[Gender == "Male"]$Age), 2)`, sd: `r round(sd(dt_included[Gender == "Male"]$Age), 2)`)
- Female: `r nrow(dt_included[Gender == "Female"])` (mean age: `r round(mean(dt_included[Gender == "Female"]$Age), 2)`, sd: `r round(sd(dt_included[Gender == "Female"]$Age), 2)`)
- Average training years: `r round(mean(dt_included$PianoTotalPractice), 2)` years (sd: `r round(sd(dt_included$PianoTotalPractice), 2)`)
- Handedness: Right (`r nrow(dt_included[Handedness == "Right"])`), Left (`r nrow(dt_included[Handedness == "Left"])`), Both (`r nrow(dt_included[Handedness == "Both"])`)
- Teaching experience (piano): Yes (`r nrow(dt_included[TeachingPiano == "Yes"])`, mean years of teaching: `r round( mean(dt_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)`, sd: `r round(sd(dt_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)`), No (`r nrow(dt_included[TeachingPiano == "No"])`)

## Some descriptive questions

1. Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.

```{r 2, echo = FALSE}
data.table("Answer" = dt_included$`Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.`)
```

```{r export, include = FALSE}
knitr::purl("questionnaire.Rmd")
```
