---
title: 'IOI-Onset Analysis for Non-Expressive performances'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

Description: This is a rough overview of non-expressive performance (IOI-Onset) from [music-teaching (experiment2)](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-2).

Non-expressive performances: data from practice trials

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}

# ggplots
theme_set(theme_classic())
theme_update(legend.position = "bottom")
```

```{r extract, include = FALSE}
dt_ioi <- fread("./preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv

# SubNr as a factor
dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)

# assign RowNr
dt_ioi$RowNr <- rep(1:72, nrow(dt_ioi)/72)
```

```{r normIOIs, include = FALSE}
# calculate normIOIs
dt_ioi$IOI <- diff(c(0, dt_ioi$TimeStamp))

# convert bpm to ms
dt_ioi[Tempo == 120]$Tempo <- 250
dt_ioi[Tempo == 110]$Tempo <- 273
dt_ioi[Tempo == 100]$Tempo <- 300

# normalise IOIs
dt_ioi$normIOI <- dt_ioi$IOI/dt_ioi$Tempo

# remove the first note
dt_ioi <- dt_ioi[RowNr != 1]

# assign Interval
dt_ioi$Interval <- rep(1:71, nrow(dt_ioi)/71)

# remove irrelevant intervals
ls_irrelevant <- list(c(25:28), c(32:35), c(51:52), c(56:57), c(65:71))
dt_ioi$EighthNote <- "Yes"

for (phrase in 1:length(ls_irrelevant)){
  for (note in 1:length(ls_irrelevant[[phrase]])){
    dt_ioi[Interval == ls_irrelevant[[phrase]][note]]$EighthNote <- "No"
  }
}
```

# IOI
## 1. All - individual
- Average normIOIs for each participant

```{r ioi, echo = FALSE}
# For each individual
ioi <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr)]
print(ioi)
```

### Plot
#### Box plot
- A dashed line represents the ideal tempo

```{r ioi_bar,  echo = FALSE}
ioi_trial <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, TrialNr)]

p_ioi <- ggplot(data = ioi_trial, aes(x = SubNr, y = Mean)) +
  geom_boxplot() +
  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5) +
  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
  labs(y = "Normalised IOI", subtitle = "IOI: Non-Expression")
p_ioi
```

#### Sequence plot
- Right wrap: Subject Number
- A dashed line represents the ideal tempo

```{r ioi_seq, fig.width = 14, fig.height = 12, echo = FALSE}
# For each individual
ioi_seq <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI)), by = .(SubNr, TrialNr, Interval)]

p_ioi_seq <- ggplot(data = ioi_seq, aes(x = Interval, y = Mean, group = TrialNr)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
  facet_grid(SubNr ~ .) +
  labs(x = "Interval", y = "Normalised IOI") + scale_x_continuous(breaks=seq(1,71,1))
p_ioi_seq
```

#### Distribution plot
- Right wrap: Trial Number
- Upper wrap: Subject Number

```{r ioi_dis, fig.width = 14, fig.height = 8, echo = FALSE}
p_ioi_dis <- ggplot(dt_ioi[EighthNote == "Yes"], aes(x = normIOI)) +
  geom_histogram(binwidth = .02, alpha = .5, position = "identity") +
  facet_grid(TrialNr ~ SubNr) +
  labs(x = "Normalised IOI", y = "Count", subtitle = "IOI: Non-Expression")
p_ioi_dis
```

## 2. All - group
- Average normIOIs

```{r ioi_all, echo = FALSE}
# Group mean
ioi_all <- ioi[, .(N = length(Mean), Mean = mean(Mean), SD = sd(Mean))]
print(ioi_all)
```

# Valid performances (exclude IOIs outliers)
- Exclude mean IOIs outside 3SD

```{r outlier, echo = FALSE}
excluded_ioi <- ioi_trial[Mean < ioi_all$Mean+3*ioi_all$SD & Mean > ioi_all$Mean-3*ioi_all$SD]
print(excluded_ioi)
```
- The number of excluded performance: `r nrow(ioi_trial)-nrow(excluded_ioi)`

# IOI Variability
## 1. All - individual
- Average CV for each participant

```{r cv, echo = FALSE}
# for each participant
cv <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), CV = sd(normIOI)/mean(normIOI)) , by = .(SubNr)]
cv
```

#### Box plot
```{r cv_box, echo = FALSE}
cv_trial <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), CV = sd(normIOI)/mean(normIOI)) , by = .(SubNr, TrialNr)]

p_cv_box <- ggplot(cv_trial, aes(x = SubNr, y = CV)) +
  geom_boxplot() +
  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5)
p_cv_box
```
## 2. All - group
- Average CV

```{r cv_all, echo = FALSE}
# for all
cv_all <- cv[, .(N = length(CV), Mean = mean(CV), SD = sd(CV))]
print(cv_all)
```

# Valid performances (exclude CVs outliers)
- Exclude mean CVs outside 3SD

```{r cv_outlier, echo = FALSE}
excluded_cv <- cv_trial[CV < cv_all$Mean+3*(cv_all$SD) & CV > cv_all$Mean-3*cv_all$SD]
print(excluded_cv)
```
- The number of excluded performance: `r cv_all$N-nrow(excluded_cv)`

```{r export, include = FALSE}
fwrite(excluded_cv, "./onset_valid.csv")
```
