---
title: "Preprocessing: trimming"
output: html_notebook
---

```{r setup, include = FALSE}
# set chunk option
knitr::opts_chunk$set(echo = FALSE)

# set working directory
if (!require("here")) {install.packages("here"); require("here")}
here::i_am("trimming.Rmd")

# packages
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}

# create a folder if not exists
if (!file.exists(here("trimmed"))){
  dir.create(here("trimmed"))
}

# read txt files
dt_onset <- fread(here("filtered", "dt_correct_onset.txt"))
dt_offset <- fread(here("filtered", "dt_correct_offset.txt"))

# sort data
setorder(dt_onset, cols = "SubNr", "TrialNr", "RowNr")
setorder(dt_offset, cols = "SubNr", "TrialNr", "RowNr")
```

```{r, include = FALSE}
# define subcomponents
# for intervals (Legato/Forte, Staccato/Piano)
ls_legato <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_staccato <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))

# for each note (velocity)
ls_legato2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_staccato2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))

# define subcomponent change (LtoS, FtoP)
change_1 <- c(5, 17, 47)
# define subcomponent change (StoL, PtoF)
change_2 <- c(8, 20, 39)
```

# IOIs
```{r, include = FALSE}
# calculate IOIs
dt_ioi <- dt_onset
dt_ioi$IOI <- diff(c(0, dt_ioi$TimeStamp))

# remove the first note
dt_ioi <- dt_ioi[RowNr != 1]

# assign Interval
dt_ioi$Interval <- rep(1:71, nrow(dt_ioi)/71)

# assign Articulation/Dynamics
dt_ioi$Subcomponent1 <- "NA"
dt_ioi$Subcomponent2 <- "NA"
# Legato/Forte
for (phrase in 1:length(ls_legato)){
  for (note in 1:length(ls_legato[[phrase]])){
    dt_ioi[Interval == ls_legato[[phrase]][note]]$Subcomponent1 <- "Legato"
    dt_ioi[Interval == ls_legato[[phrase]][note]]$Subcomponent2 <- "Forte"
  }
}
# Staccato/Piano
for (phrase in 1:length(ls_staccato)){
  for (note in 1:length(ls_staccato[[phrase]])){
    dt_ioi[Interval == ls_staccato[[phrase]][note]]$Subcomponent1 <- "Staccato"
    dt_ioi[Interval == ls_staccato[[phrase]][note]]$Subcomponent2 <- "Piano"
  }
}

# assign Subcomponent Change
for (number in change_1){
  dt_ioi[Interval == number]$Subcomponent1 <- "LtoS"
  dt_ioi[ Interval == number]$Subcomponent2 <- "FtoP"
}
for (number in change_2){
  dt_ioi[Interval == number]$Subcomponent1 <- "StoL"
  dt_ioi[Interval == number]$Subcomponent2 <- "PtoF"
}
```

## Histogram
### Before removing outliers
```{r}
dt_ioi_subset <- dt_ioi[Subcomponent1 != "NA" & !is.na(IOI)]

gghistogram(dt_ioi_subset, x = "IOI", color = "Subcomponent1", bins = 150)
```

### After removing outliers
Remove outliers, which are outside of +/- 3SD separately depending on whether responses are on boundaries (transition points between Legato/Forte and Staccato/Piano) because IOIs are likely to be slower in the transition points.

```{r}
ioi_summary <- dt_ioi_subset[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr)]

# exclude tempo deviated participants
ioi_summary$Include <- "No"
ioi_summary[Mean < mean(ioi_summary$Mean)+3*sd(ioi_summary$Mean) & Mean > mean(ioi_summary$Mean)-3*sd(ioi_summary$Mean)]$Include <- "Yes"

# evaluate if it's on boundary
dt_ioi_subset$Boundary <- "No"
dt_ioi_subset[Subcomponent1 == "LtoS" | Subcomponent1 == "StoL"]$Boundary <- "Yes"

# exclude ioi > +- 3SD (separately for each Boundary)
ioi_boundary <- dt_ioi_subset[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = Boundary]

dt_ioi_trimmed <- data.table()
for (boundary in unique(ioi_boundary$Boundary)){
  upper_ioi <- ioi_boundary[Boundary == boundary]$Mean+3*ioi_boundary[Boundary == boundary]$SD
  lower_ioi <- ioi_boundary[Boundary == boundary]$Mean-3*ioi_boundary[Boundary == boundary]$SD
  dt_current <- dt_ioi_subset[Boundary == boundary & IOI < upper_ioi & IOI > lower_ioi]
  dt_ioi_trimmed <- rbind(dt_ioi_trimmed, dt_current)
}

gghistogram(dt_ioi_trimmed, color = "Subcomponent1", x = "IOI", bins = 150)
```

Remove `r nrow(dt_ioi_subset)-nrow(dt_ioi_trimmed)` responses beyond +- 3SD / `r 100*((nrow(dt_ioi_subset)-nrow(dt_ioi_trimmed))/nrow(dt_ioi_subset))` % of total responses

```{r, include = FALSE}
fwrite(dt_ioi_trimmed, file = here("trimmed", "dt_ioi_trimmed.txt"), row.names = F)
```

# KOT
```{r, include = FALSE}
dt_kot <- dt_onset

# calculate KOT (Offset 1 - Onset 2)
dt_kot$KOT <- NA
for (row in 1:nrow(dt_kot)){
  if (row < nrow(dt_kot)){
    dt_kot$KOT[row+1] <- dt_offset$TimeStamp[row] - dt_onset$TimeStamp[row+1] # offset(n) - onset(n+1)
  }
}

# remove the first note
dt_kot <- dt_kot[RowNr != 1]

# assign a sequence number for each tone
dt_kot$Interval <- rep(1:71, nrow(dt_kot)/71)

# assign Articulation/Dynamics
dt_kot$Subcomponent1 <- "NA"
dt_kot$Subcomponent2 <- "NA"
# Legato/Forte
for (phrase in 1:length(ls_legato)){
  for (note in 1:length(ls_legato[[phrase]])){
    dt_kot[Interval == ls_legato[[phrase]][note]]$Subcomponent1 <- "Legato"
    dt_kot[Interval == ls_legato[[phrase]][note]]$Subcomponent2 <- "Forte"
  }
}
# Staccato/Piano
for (phrase in 1:length(ls_staccato)){
  for (note in 1:length(ls_staccato[[phrase]])){
    dt_kot[Interval == ls_staccato[[phrase]][note]]$Subcomponent1 <- "Staccato"
    dt_kot[Interval == ls_staccato[[phrase]][note]]$Subcomponent2 <- "Piano"
  }
}

# assign Subcomponent Change
for (number in change_1){
  dt_kot[Interval == number]$Subcomponent1 <- "LtoS"
  dt_kot[ Interval == number]$Subcomponent2 <- "FtoP"
}
for (number in change_2){
  dt_kot[Interval == number]$Subcomponent1 <- "StoL"
  dt_kot[Interval == number]$Subcomponent2 <- "PtoF"
}
```

## Histogram
### Before removing outliers

```{r}
dt_kot_subset <- dt_kot[Subcomponent1 != "NA" & !is.na(KOT)]

gghistogram(dt_kot_subset, x = "KOT", color = "Subcomponent1", bins = 150)
```

### After removing outliers
```{r}
kot_subcomponent <- dt_kot_subset[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = Subcomponent1]
dt_kot_trimmed <- data.table()
for (subcomponent in unique(dt_kot_subset$Subcomponent1)){
  upper <- kot_subcomponent[Subcomponent1 == subcomponent]$Mean+3*kot_subcomponent[Subcomponent1 == subcomponent]$SD
  lower <- kot_subcomponent[Subcomponent1 == subcomponent]$Mean-3*kot_subcomponent[Subcomponent1 == subcomponent]$SD
  dt_current <- dt_kot_subset[Subcomponent1 == subcomponent & KOT < upper & KOT > lower]
  dt_kot_trimmed <- rbind(dt_kot_trimmed, dt_current)
}

gghistogram(dt_kot_trimmed, x = "KOT", color = "Subcomponent1", bins = 150)
```

Remove `r nrow(dt_kot_subset)-nrow(dt_kot_trimmed)` responses beyond +- 3SD / `r 100*((nrow(dt_kot_subset)-nrow(dt_kot_trimmed))/nrow(dt_kot_subset))` % of total responses

```{r, include = FALSE}
fwrite(dt_kot_trimmed, file = here("trimmed", "dt_kot_trimmed.txt"), row.names = F)
```

# KV
```{r, include = FALSE}
# calculate Diff (acceleration - velocity difference between notes)
dt_vel <- dt_onset
dt_vel$Diff <- diff(c(0, dt_vel$Velocity))

# remove the first note
dt_vel_diff <- dt_vel[RowNr != 1]
dt_vel$Diff <- NULL # remove Diff from dt_vel

# assign Subcomponents
# for each note
dt_vel$Subcomponent1 <- "NA"
dt_vel$Subcomponent2 <- "NA"
# Legato/Forte
for (phrase in 1:length(ls_legato2)){
  for (note in 1:length(ls_legato2[[phrase]])){
    dt_vel[RowNr == ls_legato2[[phrase]][note]]$Subcomponent1 <- "Legato"
    dt_vel[RowNr  == ls_legato2[[phrase]][note]]$Subcomponent2 <- "Forte"
  }
}
# Staccato/Piano
for (phrase in 1:length(ls_staccato2)){
  for (note in 1:length(ls_staccato2[[phrase]])){
    dt_vel[RowNr  == ls_staccato2[[phrase]][note]]$Subcomponent1 <- "Staccato"
    dt_vel[RowNr  == ls_staccato2[[phrase]][note]]$Subcomponent2 <- "Piano"
  }
}
```

## Histogram
### Before removing outliers
```{r}
dt_vel_subset <- dt_vel[Subcomponent2 != "NA" & !is.na(Velocity)]

gghistogram(dt_vel_subset, x = "Velocity", color = "Subcomponent2", bins = 150)
```

### After removing outliers
```{r}
vel_subcomponent <- dt_vel_subset[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = Subcomponent2]
dt_vel_trimmed <- data.table()
for (subcomponent in unique(dt_vel_subset$Subcomponent2)){
  upper <- vel_subcomponent[Subcomponent2 == subcomponent]$Mean+3*vel_subcomponent[Subcomponent2 == subcomponent]$SD
  lower <- vel_subcomponent[Subcomponent2 == subcomponent]$Mean-3*vel_subcomponent[Subcomponent2 == subcomponent]$SD
  dt_current <- dt_vel_subset[Subcomponent2 == subcomponent & Velocity < upper & Velocity > lower]
  dt_vel_trimmed <- rbind(dt_vel_trimmed, dt_current)
}

gghistogram(dt_vel_trimmed, x = "Velocity", color = "Subcomponent2", bins = 150)
```

Remove `r nrow(dt_vel_subset)-nrow(dt_vel_trimmed)` responses beyond +- 3SD / `r 100*((nrow(dt_vel_subset)-nrow(dt_vel_trimmed))/nrow(dt_vel_subset))` % of total responses

```{r, include = FALSE}
fwrite(dt_vel_trimmed, file = here("trimmed", "dt_vel_trimmed.txt"), row.names = F)
```

# KV Difference
```{r, include = FALSE}
# assign Interval
dt_vel_diff$Interval <- rep(1:71, nrow(dt_vel_diff)/71)

# assign Subcomponents
# for intervals
dt_vel_diff$Subcomponent1 <- "NA"
dt_vel_diff$Subcomponent2 <- "NA"
# Legato/Forte
for (phrase in 1:length(ls_legato)){
  for (note in 1:length(ls_legato[[phrase]])){
    dt_vel_diff[Interval == ls_legato[[phrase]][note]]$Subcomponent1 <- "Legato"
    dt_vel_diff[Interval == ls_legato[[phrase]][note]]$Subcomponent2 <- "Forte"
  }
}
# Staccato/Piano
for (phrase in 1:length(ls_staccato)){
  for (note in 1:length(ls_staccato[[phrase]])){
    dt_vel_diff[Interval == ls_staccato[[phrase]][note]]$Subcomponent1 <- "Staccato"
    dt_vel_diff[Interval == ls_staccato[[phrase]][note]]$Subcomponent2 <- "Piano"
  }
}

# assign Subcomponent Change
for (number in change_1){
  dt_vel_diff[Interval == number]$Subcomponent1 <- "LtoS"
  dt_vel_diff[ Interval == number]$Subcomponent2 <- "FtoP"
}
for (number in change_2){
  dt_vel_diff[Interval == number]$Subcomponent1 <- "StoL"
  dt_vel_diff[Interval == number]$Subcomponent2 <- "PtoF"
}

```

## Histogram
### Before removing outliers
```{r}
dt_vel_diff_subset <- dt_vel_diff[Subcomponent2 != "NA" & !is.na(Diff)]

gghistogram(dt_vel_diff_subset, x = "Diff", color = "Subcomponent2", bins = 150)
```

### After removing outliers
```{r}
vel_diff_subcomponent <- dt_vel_diff_subset[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = Subcomponent2]
dt_vel_diff_trimmed <- data.table()
for (subcomponent in unique(dt_vel_diff_subset$Subcomponent2)){
  upper <- vel_diff_subcomponent[Subcomponent2 == subcomponent]$Mean+3*vel_diff_subcomponent[Subcomponent2 == subcomponent]$SD
  lower <- vel_diff_subcomponent[Subcomponent2 == subcomponent]$Mean-3*vel_diff_subcomponent[Subcomponent2 == subcomponent]$SD
  dt_current <- dt_vel_diff_subset[Subcomponent2 == subcomponent & Diff < upper & Diff > lower]
  dt_vel_diff_trimmed <- rbind(dt_vel_diff_trimmed, dt_current)
}

gghistogram(dt_vel_diff_trimmed, x = "Diff", color = "Subcomponent2", bins = 150)
```

Remove `r nrow(dt_vel_diff_subset)-nrow(dt_vel_diff_trimmed)` responses beyond +- 3SD / `r 100*((nrow(dt_vel_diff_subset)-nrow(dt_vel_diff_trimmed))/nrow(dt_vel_diff_subset))` % of total responses

```{r, include = FALSE}
fwrite(dt_vel_diff_trimmed, file = here("trimmed", "dt_vel_diff_trimmed.txt"), row.names = F)
```
