---
title: 'IOI-Onset Selection'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: In order to select 4 performances to determine a tempo of each performance, we look at previously generated 16 instances and randomly selected from them.

Original stimuli: `material/expstimuli/original/averaging/low/1596207520-310720/dt_ioi_instance.txt`

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r extract, include = FALSE}
dt_ioi <- fread("../../original/averaging/low/1596207520-310720/dt_ioi_instance.txt")
dt_ioi$IOI <- dt_ioi$Mean*300
```

# IOIs plot
```{r plot-ioi, echo = FALSE, fig.height = 6}
ggplot(data = dt_ioi, aes(x = RowNr, y = IOI)) +
  geom_line() +
  geom_point() +
  facet_grid(Instance ~ .) +
  labs(x = "RowNr", y = "IOIs (100bpm)", subtitle = "IOI: 16 instances") +
  scale_x_continuous(breaks=seq(1,72,1))
```
- Looks like 13 may have a delay at Row 24?

```{r playback, include = FALSE}
# create a folder to store stimuli
foldername = paste(format(Sys.time(), "%s-%d%m%y"), "/", sep = "") # current time
dir.create(foldername)

# 1. determine onsets/offsets
# onsets
dt_playback_onset <- dt_ioi[, c("Instance", "RowNr")]
dt_playback_onset$Key_OnOff <- 1
for (i in 1:nrow(dt_playback_onset)){
  if (dt_playback_onset$RowNr[i] == 1){
    dt_playback_onset$TimeStamp[i] <- 0
  } else {
    dt_playback_onset$TimeStamp[i] <- dt_playback_onset$TimeStamp[i-1]+dt_ioi$Mean[i]
  }
}
# Tempo 100 bpm (IOI - 300ms (8th notes))
dt_playback_onset$TimeStamp <- round(dt_playback_onset$TimeStamp*300)

# offsets
dt_playback_offset <- dt_ioi[, c("Instance", "RowNr")]
dt_playback_offset$Key_OnOff <- 0
# Tempo 100 bpm (IOI - 300ms (8th notes))
dt_playback_offset$TimeStamp <- round(dt_playback_onset$TimeStamp+0.5*300)

# kv
dt_playback_onset$Velocity <- 80
dt_playback_offset$Velocity <- 64

# create txt for each instance
for (i in 1:16){
  onset <- dt_playback_onset[Instance == i]
  offset <- dt_playback_offset[Instance == i]
  onset$Pitch <- 60
  offset$Pitch <- 60
  instance <- rbind(onset, offset)
  # sort by TimeStamp
  instance <- instance[order(TimeStamp)]
  
  # export txt
  filename = paste(foldername, i, "_instance.txt", sep = "")
  fwrite(instance, filename)
}
```

# Export as mid files
```{bash, echo = TRUE}
# python
source /Users/atsukotominaga/Pyenv/bin/activate
python mid_export.py
```
- Listened to each performance (mid files between 1-16) and `ioi_13.mid` was excluded because there was one tiny delay in a performance.
- We select four of these performances (15 mid files) to determine a tempo of each stimulus.

# Selected 4 performances (IOIs)
```{r, echo = FALSE}
select_ioi <- dt_ioi[Instance != 13, .N, .(Instance)]
select_ioi$Sample <- sample(c(1:nrow(select_ioi)), replace = FALSE)
select_ioi[Sample < 5]

# export
dir.create("./tempo/")
for (i in 1:4){
  current <- dt_ioi[Instance == select_ioi[Sample == i]$Instance]
  filename = paste("./tempo/", i, ".txt", sep = "")
  fwrite(current, filename)
}
```
